---
title: "Data_SetUp"
output: html_document
date: "2025-03-04"
---

We begin with the first part of our project, exploring the many features involved in our dataset. In particular, we will probe into the specifics, diving into features for our different trials, sessions, and mice. This will include session-specific data structures, neural activities occurring in each trial, changes across trials, and the similarities and differences between sessions and between mice.

After this analysis, we should have a better understanding of our dataset and its several features. Thus, our evolved understanding will allow for a more representative predictive model.

```{r}
# """Bringing all of our data together"""


session=list()
for(i in 1:18){
  session[[i]]=readRDS(paste('./Data/session',i,'.rds',sep=''))
  
   print(session[[i]]$mouse_name)
   print(session[[i]]$date_exp)
}
```

We can visualize some of our data using the table produced below. In this table, we will extract session information -- this will include the date of the experiment, the number of brain areas that were involved, the number of neurons that were activated, the number of trials, and the success rate; All of this data will be collected for each mouse.

```{r}
library(tidyverse)
library(knitr)
library(xtable)
library(kableExtra)

n.session = length(session)
meta <- tibble(
  mouse_name = rep('name',n.session),
  date_exp =rep('dt',n.session),
  n_brain_area = rep(0,n.session),
  n_neurons = rep(0,n.session),
  n_trials = rep(0,n.session),
  success_rate = rep(0,n.session)
)

for(i in 1:n.session){
  tmp = session[[i]];
  meta[i,1]=tmp$mouse_name;
  meta[i,2]=tmp$date_exp;
  meta[i,3]=length(unique(tmp$brain_area));
  meta[i,4]=dim(tmp$spks[[1]])[1];
  meta[i,5]=length(tmp$feedback_type);
  meta[i,6]=mean(tmp$feedback_type+1)/2;
  }


meta %>%
    kbl(digits=2, col.names = c("Mice", "Experiment Date", "Number of Brain Areas", "Number of Neurons", "Number of Trials", "Success Rate")) %>%
    kable_classic(full_width=F, html_font="Cambria")
```
As we can see from the table above, the data per mouse exhibits considerable heterogeneity. 

For instance, while Hench and Forssmann were both subjected to four sessions of experimentation, Cori only underwent three, while Lederberg underwent seven sessions. This inequity in experimentation may impact the substance of our data by making our information more representative of our heavily-tested specimen, Lederberg.




Here, we reflect on the differences within our study.

```{r}


contrasts <- tibble(
  mouse_name = rep('name',n.session),
  date_exp = rep('df',n.session),
  n_brain_area = rep(0,n.session),
  n_neurons = rep(0,n.session),
  n_trials = rep(0,n.session),
  success_rate = rep(0,n.session)
)
View(meta)

amount=0

for(i in 1:n.session){
  tmp = session[[i]];
  contrasts[i,1]=tmp$mouse_name;
  contrasts[i,2]=tmp$date_exp
  contrasts[i,3]=length(unique(tmp$brain_area));
  contrasts[i,4]=dim(tmp$spks[[1]])[1];
  contrasts[i,5]=length(tmp$feedback_type);
  contrasts[i,6]=mean(tmp$feedback_type+1)/2;
  }
View(contrasts)

#A function to get the date_exp
alter_Date <- function(data, mName, func){
  theResult = func((data %>% filter(mouse_name == mName))$date_exp)
  return (theResult)
}
#A function to get n_trials
alter_nTrials <- function(data, mName, func){
  theResult = func((data %>% filter(mouse_name == mName))$n_trials)
  return (theResult)
}
#A function to get success_rate
alter_nTrials <- function(data, mName, func){
  theResult = func((data %>% filter(mouse_name == mName))$n_trials)
  return (theResult)
}
```

```{r}
get_Data <- function(data, mName, var, minMax){
  if (var=="success_rate"){
    answer = ((data %>% filter(mouse_name == mName))$success_rate)
  }else if (var=="n_brain_area"){
    answer = ((data %>% filter(mouse_name == mName))$n_brain_area)
  } else {
    answer = ((data %>% filter(mouse_name == mName))$n_neurons)
  }
  if (minMax == 0){
    answer = answer[1]
  } else {
    dataLen = length(data)
    answer = tail(answer, 1)
  }
  answer = signif(answer, 2)
  return (answer)
}

getMeans <- function(mName, var){
  sum = 0
  count = 0
  for (i in 1:n.session){
    tmp = session[[i]];
    if (tmp$mouse_name == mName){
      if (var == "success_rate"){
         val = mean(tmp$feedback_type+1)/2
         sum = sum+val
      } else if (var == "n_brain_area"){
         val = length(unique(tmp$brain_area))
         sum = sum+val
      }else {
         val = dim(tmp$spks[[1]])[1]
         sum = sum+val
      }
      count = count+1
    }
  }
  meanSuccess = (sum/count)
  if (var == "n_brain_area"){
    meanSuccess = signif((sum/count),1)
  }
  return (meanSuccess)
}

```


```{r}
#Find the number of sessions for each mouse
cor.exps = alter_Date(contrasts, "Cori", length);
for.exps = alter_Date(contrasts, "Forssmann", length); 
hen.exps = alter_Date(contrasts, "Hench", length); 


#Find the number of trials for each mouse
cor.trials = alter_nTrials(contrasts, "Cori", sum); 
for.trials = alter_nTrials(contrasts, "Forssmann", sum); 
hen.trials = alter_nTrials(contrasts, "Hench", sum);


#First, last, and mean data points, for specified data

#Cori 
cor.success_rate.mean = getMeans("Cori", "success_rate")
cor.success_rate.first = get_Data(contrasts, "Cori", "success_rate", 0)
cor.success_rate.last = get_Data(contrasts, "Cori", "success_rate", 1)


cor.n_brain_area.mean = getMeans("Cori", "n_brain_area")
cor.n_brain_area.first = get_Data(contrasts, "Cori", "n_brain_area", 0)
cor.n_brain_area.last = get_Data(contrasts, "Cori", "n_brain_area", 1)


cor.n_neurons.mean = getMeans("Cori", "n_neurons")
cor.n_neurons.first = get_Data(contrasts, "Cori", "n_neurons", 0)
cor.n_neurons.last = get_Data(contrasts, "Cori", "n_neurons", 1)



#Forssmann
for.success_rate.mean = getMeans("Forssmann", "success_rate")
for.success_rate.first = get_Data(contrasts, "Forssmann", "success_rate", 0)
for.success_rate.last = get_Data(contrasts, "Forssmann", "success_rate", 1)


for.n_brain_area.mean = getMeans("Forssmann", "n_brain_area")
for.n_brain_area.first = get_Data(contrasts, "Forssmann", "n_brain_area", 0)
for.n_brain_area.last = get_Data(contrasts, "Forssmann", "n_brain_area", 1)


for.n_neurons.mean = getMeans("Forssmann", "n_neurons")
for.n_neurons.first = get_Data(contrasts, "Forssmann", "n_neurons", 0)
for.n_neurons.last = get_Data(contrasts, "Forssmann", "n_neurons", 1)



#Hench
hen.success_rate.mean = getMeans("Hench", "success_rate")
hen.success_rate.first = get_Data(contrasts, "Hench", "success_rate", 0)
hen.success_rate.last = get_Data(contrasts, "Hench", "success_rate", 1)


hen.n_brain_area.mean = getMeans("Hench", "n_brain_area")
hen.n_brain_area.first = get_Data(contrasts, "Hench", "n_brain_area", 0)
hen.n_brain_area.last = get_Data(contrasts, "Hench", "n_brain_area", 1)


hen.n_neurons.mean = getMeans("Hench", "n_neurons")
hen.n_neurons.first = get_Data(contrasts, "Hench", "n_neurons", 0)
hen.n_neurons.last = get_Data(contrasts, "Hench", "n_neurons", 1)



#Lederberg
led.success_rate.mean = getMeans("Lederberg", "success_rate")
led.success_rate.first = get_Data(contrasts, "Lederberg", "success_rate", 0)
led.success_rate.last = get_Data(contrasts, "Lederberg", "success_rate", 1)


led.n_brain_area.mean = getMeans("Lederberg", "n_brain_area")
led.n_brain_area.first = get_Data(contrasts, "Lederberg", "n_brain_area", 0)
led.n_brain_area.last = get_Data(contrasts, "Lederberg", "n_brain_area", 1)


led.n_neurons.mean = getMeans("Lederberg", "n_neurons")
led.n_neurons.first = get_Data(contrasts, "Lederberg", "n_neurons", 0)
led.n_neurons.last = get_Data(contrasts, "Lederberg", "n_neurons", 1)

```


```{r}
#######Ongoing
#Table for the first, last, and mean success rate
n.session = length(session)
success_rate.Data <- tibble(
  mouse_name = rep(0,4),
  min = rep(0,4),
  max = rep(0,4),
  mean = rep(0,4),
)


for(i in range(4)){
  success_rate.Data$mouse_name = c("Cori", "Forssmann", "Hench", "Lederberg")
  success_rate.Data$min = c(cor.success_rate.first, for.success_rate.first, hen.success_rate.first, led.success_rate.first)
  success_rate.Data$max = c(cor.success_rate.last, for.success_rate.last, hen.success_rate.last, led.success_rate.last)
  success_rate.Data$mean = c(cor.success_rate.mean, for.success_rate.mean, hen.success_rate.mean, led.success_rate.mean)
}
success_rate.Data %>%
    kbl(digits=2, col.names = c("Mice", "First Success Rate", "Last Success Rate", "Mean Success Rate")) %>%
    kable_classic(full_width=F, html_font="Cambria")

#Table for the first, last, and mean number of brain areas
n_brain_area.Data <- tibble(
  mouse_name = rep(0,4),
  min = rep(0,4),
  max = rep(0,4),
  mean = rep(0,4),
)


for(i in range(4)){
  n_brain_area.Data$mouse_name = c("Cori", "Forssmann", "Hench", "Lederberg")
  n_brain_area.Data$min = c(cor.n_brain_area.first, for.n_brain_area.first, hen.n_brain_area.first, led.n_brain_area.first)
  n_brain_area.Data$max = c(cor.n_brain_area.last, for.n_brain_area.last, hen.n_brain_area.last, led.n_brain_area.last)
  n_brain_area.Data$mean = c(cor.n_brain_area.mean, for.n_brain_area.mean, hen.n_brain_area.mean, led.n_brain_area.mean)
}
n_brain_area.Data %>%
    kbl(digits=2, col.names = c("Mice", "First Number of Brain Areas", "Last Number of Brain Areas", "Mean Number of Brain Areas")) %>%
    kable_classic(full_width=F, html_font="Cambria")


#Table for the first, last, and mean number of neurons
n_neurons.Data <- tibble(
  mouse_name = rep(0,4),
  min = rep(0,4),
  max = rep(0,4),
  mean = rep(0,4),
)


for(i in range(4)){
  n_neurons.Data$mouse_name = c("Cori", "Forssmann", "Hench", "Lederberg")
  n_neurons.Data$min = c(cor.n_neurons.first, for.n_neurons.first, hen.n_neurons.first, led.n_neurons.first)
  n_neurons.Data$max = c(cor.n_neurons.last, for.n_neurons.last, hen.n_neurons.last, led.n_neurons.last)
  n_neurons.Data$mean = c(cor.n_neurons.mean, for.n_neurons.mean, hen.n_neurons.mean, led.n_neurons.mean)
}
n_neurons.Data %>%
    kbl(digits=2, col.names = c("Mice", "First Number of Neurons", "Last Number of Neurons", "Mean Number of Neurons")) %>%
    kable_classic(full_width=F, html_font="Cambria")



cor.exps = alter_Date(contrasts, "Cori", length); cor.exps
for.exps = alter_Date(contrasts, "Forssmann", length); for.exps
hen.exps = alter_Date(contrasts, "Hench", length); hen.exps
led.exps = alter_Date(contrasts, "Lederberg", length); led.exps

#Find the number of trials for each mouse
cor.trials = alter_nTrials(contrasts, "Cori", sum); cor.trials
for.trials = alter_nTrials(contrasts, "Forssmann", sum); for.trials
hen.trials = alter_nTrials(contrasts, "Hench", sum); hen.trials
led.trials = alter_nTrials(contrasts, "Lederberg", sum); led.trials


#####Ongoing
generalSummary.Data <- tibble(
  mouse_name = rep(0,4),
  n.sessions = rep(0,4),
  n.trials = rep(0,4),
  success_rate.min = rep(0,4),
  success_rate.max = rep(0,4),
  success_rate.mean = rep(0,4),
  n_brain_area.min = rep(0,4),
  n_brain_area.max = rep(0,4),
  n_brain_area.mean = rep(0,4),
  n_neurons.min = rep(0,4),
  n_neurons.max = rep(0,4),
  n_neurons.mean = rep(0,4),
)

for(i in range(4)){
  generalSummary.Data$mouse_name = c("Cori", "Forssmann", "Hench", "Lederberg")
  generalSummary.Data$n.sessions = c(cor.exps, for.exps, hen.exps, led.exps)
  generalSummary.Data$n.trials = c(cor.trials, for.trials, hen.trials, led.trials)
  generalSummary.Data$min = c(cor.success_rate.first, for.success_rate.first, hen.success_rate.first, led.success_rate.first)
  generalSummary.Data$max = c(cor.success_rate.last, for.success_rate.last, hen.success_rate.last, led.success_rate.last)
  generalSummary.Data$mean = c(cor.success_rate.mean, for.success_rate.mean, hen.success_rate.mean, led.success_rate.mean)
  generalSummary.Data$min = c(cor.n_brain_area.first, for.n_brain_area.first, hen.n_brain_area.first, led.n_brain_area.first)
  generalSummary.Data$max = c(cor.n_brain_area.last, for.n_brain_area.last, hen.n_brain_area.last, led.n_brain_area.last)
  generalSummary.Data$mean = c(cor.n_brain_area.mean, for.n_brain_area.mean, hen.n_brain_area.mean, led.n_brain_area.mean)
  generalSummary.Data$min = c(cor.n_neurons.first, for.n_neurons.first, hen.n_neurons.first, led.n_neurons.first)
  generalSummary.Data$max = c(cor.n_neurons.last, for.n_neurons.last, hen.n_neurons.last, led.n_neurons.last)
  generalSummary.Data$mean = c(cor.n_neurons.mean, for.n_neurons.mean, hen.n_neurons.mean, led.n_neurons.mean)
}



# generalSummary.Data %>%
#     kbl(col.names = c("Mice", "Number of Session", "Number of Trials", "First Success Rate", "Last Success Rate", "Mean Success Rate", "First Number of Brain Areas", "Last Number of Brain Areas", "Mean Number of Brain Areas", "First Number of Neurons", "Last Number of Neurons", "Mean Number of Neurons")) %>%
#     kable_classic(full_width=F, html_font="Cambria")
# 


```


As demonstrated in the tables above, there are significant differences between the starting, ending, and mean values for each mouse throughout this study. 
In the tables above, we studied the starting, ending, and mean values for the success rate ('success_rate'), the number of brain areas ('n_brain_area'), and the number of involved neurons ('n_neurons'). 


First, we examine the success rates. 
Not all of the success rates see a significant increase in the study--while the other three mice have an ending success rate that is larger than their first success rate, Forssmann has the same starting and ending success rate. Given that this mouse underwent the same experiment multiple times and failed to improve along with its peers, we can guess that the mouse's individual qualities had something to do with this abnormality.
Lederberg had the highest mean success rate--This can be attributed to individual qualities or (very likely) how Lederberg underwent more sessions and more trials than any other mouse in the study.

Then, we examine the number of brain areas.







To better understand our data (and the structures of our sessions), we can select a session and plot its spikes per area.
Based upon the spikes per brain area, we can identify the brain areas we believe to be most active, characterized by the largest spikes.


```{r}
average_spike_area<-function(i.t,this_session){
  spk.trial = this_session$spks[[i.t]]
  area= this_session$brain_area
  spk.count=apply(spk.trial,1,sum)
  spk.average.tapply=tapply(spk.count, area, mean)
  return(spk.average.tapply)
  }

# Test the function

i.s=17
i.t=1

average_spike_area(1,this_session = session[[i.s]])
```




```{r}
n.trial=length(session[[i.s]]$feedback_type)
n.area=length(unique(session[[i.s]]$brain_area ))
# Alternatively, you can extract these information in the meta that we created before.

# We will create a data frame that contain the average spike counts for each area, feedback type,  the two contrasts, and the trial id

trial.summary =matrix(nrow=n.trial,ncol= n.area+1+2+1)
for(i.t in 1:n.trial){
  trial.summary[i.t,]=c(average_spike_area(i.t,this_session = session[[i.s]]),
                          session[[i.s]]$feedback_type[i.t],
                        session[[i.s]]$contrast_left[i.t],
                        session[[i.s]]$contrast_right[i.s],
                        i.t)
}

colnames(trial.summary)=c(names(average_spike_area(i.t,this_session = session[[i.s]])), 'feedback', 'left contr.','right contr.','id' )

# Turning it into a data frame
trial.summary <- as_tibble(trial.summary)
```


Now, we plot our spike data.

```{r}
area.col=rainbow(n=n.area,alpha=0.7)
# In base R, I usually initiate a blank plot before drawing anything on it
plot(x=1,y=0, col='white',xlim=c(0,n.trial),ylim=c(0.5,2.2), xlab="Trials",ylab="Average spike counts", main=paste("Spikes per area in Session", i.s))


for(i in 1:n.area){
  lines(y=trial.summary[[i]],x=trial.summary$id,col=area.col[i],lty=2,lwd=1)
  lines(smooth.spline(trial.summary$id, trial.summary[[i]]),col=area.col[i],lwd=3)
  }
legend("topright", 
  legend = colnames(trial.summary)[1:n.area], 
  col = area.col, 
  lty = 1, 
  cex = 0.8
)

```

Judging by the graph above, we can see that the average spike counts fluctuate considerably. The highest average spike counts appear to be associated with brain areas "LD" and "VPL". The biggest fluctuation appeared to come from the "RT" brain area, where the average spike count jumped up by about a unit, in the area between the 80th trial and the 100th trial.

























#Sources
Source 1: For kableExtra Table, on Line 26 --> https://haozhu233.github.io/kableExtra/awesome_table_in_html.html

